Management induced changes of soil organic carbon on global croplands

### 2020-12-14 SOCbudgetPaper/conclusion.Rmd
update conclusion

The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.

>>>>

Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management. It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.

Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models. That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data. Our data-set and the MADRaT package ([@dietrich_madrat_2020]) constitute a starting point for building comprehensive data sets on agricultural management aspects.

Our study again highlights that the expansion of croplands is still a major source of CO~2~ emissions -- not only by the removal of vegetation, but also by a slow depletion of soils. Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.

However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975. Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils. The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO~2~ source. Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO~2~. The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.

Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks. Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand ([@popp_land-use_2014-1]) -- and thereby usually further reduce the SOC debt on croplands.

While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets. Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.

Yet, there is ample potential for further improved SOC management. As shown in Fig. \@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year). Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC). Improved management could include, firstly, a circular flow from the food supply chain back to soils. Waste composting or excreta recycling could represent a major additional N input to cropland soils ([@brenzinger_organic_2018]). Secondly, soil carbon sequestration techniques ([@smith_soil_2016]), deep ploughing ([@alcantara_deep_2016]) or the transformation of c inputs to more recalcitrant biochar ([@woolf_sustainable_2010]) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools. Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs. Finally, other carbon-accumulating practices, such as the cultivation of cover crops ([@poeplau_carbon_2015]) and agroforestry ([@lorenz_soil_2014]) could increase total C sequestration on croplands.
